% !TeX root = ../sustechthesis-example.tex

% 中英文摘要和关键字

\begin{abstract}
  足式机器人相对于常见的轮式和履带式机器人有着更加灵活，环境适应能力强的优势。与启发它们的人类或者动物一样，这类机器人拥有着更强的越野能力，能够适应复杂的地形环境，到达普通轮式或履带式机器人无法到达的地方。然而，这类机器人往往具有更多的自由度，这使得对它们的控制变得具有挑战性。如何让这类足式机器人像它们的动物模仿对象一样能够灵活、优雅地运动成为一个重要的研究课题。在这其中，模仿四足动物的四足机器人（机械狗）是一类被广泛研究的对象。

  机械狗是一个多自由度运动系统，其典型自由度数量为12个。每条腿有三个自由度分别实现小腿伸缩、大腿伸缩和整腿伸缩。由于它的求解空间巨大而可用的解空间相对较小，因此机械狗的运动控制常常是启发式的——预先定义一些运动，以缩小求解时的搜索空间。经典的基于运动动力学建模和优化问题的控制方式主要有\emph{轨迹优化(Trajectory Optimization, TO)}和\emph{模型预测控制(Model Predictive Control, MPC)}。由于求解需要涉及大量的非线性优化问题，这类控制方法求解算力开销较大往往不能够在线完成，需要上位机的辅助。
  相比较而言近年来发展起来的基于深度学习的\emph{强化学习控制}方法在运行期间的算力开销很小。\emph{强化学习控制}是一种通过构建神经网络并依靠大量训练优化神经网络参数以得到预期的控制效果的控制方法，可以完全不需要动力学建模。其训练过程往往在仿真环境中完成，因此需要对机械狗的物理模型进行建模，这些建模包括：机械狗的躯干及四肢的刚体模型、电机驱动关节的响应模型。

  不过由于神经网络的黑盒性，单纯的强化学习控制往往只能针对特定的训练目标而很难与其它模型融合，缺乏跨模式泛化能力。因此经典控制和强化学习控制的融合控制变得十分重要。



  % 关键词用“英文逗号”分隔，输出时会自动处理为正确的分隔符
  \thusetup{
    keywords = {机器人控制, 轨迹优化, 深度学习, 强化学习控制, Isaac仿真},
  }
\end{abstract}

% \begin{abstract*}
%   An abstract of a dissertation is a summary and extraction of research work and contributions.
%   Included in an abstract should be description of research topic and research objective, brief introduction to methodology and research process, and summarization of conclusion and contributions of the research.
%   An abstract should be characterized by independence and clarity and carry identical information with the dissertation.
%   It should be such that the general idea and major contributions of the dissertation are conveyed without reading the dissertation.

%   An abstract should be concise and to the point.
%   It is a misunderstanding to make an abstract an outline of the dissertation and words “the first chapter”, “the second chapter” and the like should be avoided in the abstract.

%   For thesis written in \textbf{Chinese} as the main text, the abstract of doctoral thesis is about 800 to 1000 words; and the abstract of the master's thesis is generally about 500 words; the length is limited to one page.
%   For thesis written in \textbf{English} as the main text, the word count of Chinese abstracts is required to be 800 to 1000 words (regardless doctoral thesis or master's thesis).

%   The abstract of the doctoral thesis is about 800 to 1000 words; the abstract of the master's thesis is generally about 500 words; the length is limited to one page.

%   Keywords are terms used in a dissertation for indexing, reflecting core information of the dissertation.
%   An abstract may contain a maximum of 5 keywords, with semi-colons used in between to separate one another.

%   % Use comma as seperator when inputting
%   \thusetup{
%     keywords* = {keyword 1, keyword 2, keyword 3, keyword 4, looooooooooooooooooooong keyword 5},
%   }
% \end{abstract*}
